### 2020-03-07 short.body.tex
k i think this color scheme is more or less ok for now

In an epidemic, symptomatic cases are the predominant focus of treatment and usually represent the bulk of reported cases. 
However, infected individuals who are asymptomatic yet infectious can be a critical factor in the spread of some pathogens \citep{fraser2004factors}.
Asymptomatic individuals are hard to trace, unlikely to self-isolate, and are likely to retain normal social and travel patterns.
The COVID-19 outbreak has raised significant
questions regarding the role of asymptomatic cases~\citep{fauci_nejm2020}.
A key focus has been on estimating asymptomatic prevalence (see~\citep{mizumoto_2020}), e.g., what fraction of individuals have asymptomatic infections relative to those with symptomatic infections?
%.(i) What fraction of newly infected individuals
%have mild and/or subclinical infections rather than severe and/or
%clinical infections? (ii) What is the time-scale over which 
%individuals who have mild and/or subclinical cases remain
%infectious? (iii) How infectious are individuals
%with mild and/or subclinical infections relative to those
%with severe symptoms?

@@

We model viral spread using a renewal-equation framework \citep{heesterbeek1996concept}, which allows us to model the current incidence of infected individuals as a function of previous incidence and how infectiousness of an infected individual varies over the course of their infection;
in Appendix, we show that results derived from a compartmental model is a special case of the result we present here.
We divide incidence $i$ into two categories -- $i_a$ and $i_s$ -- corresponding to incidence of asymptomatic and symptomatic cases, respectively.
Both types of individuals can infect others, but they differ in their intrinsic reproduction numbers, ${\cal{R}}_a$ and ${\cal{R}}_s$, as well as intrinsic generation-interval distributions, $g_a(\tau)$ and $g_s(\tau)$.
Generation intervals, which are defined as the time between when an individual is infected and when that individual infects another person \citep{svensson2007note}, depend on the natural history of infection:
individuals with subclinical cases may have fast clearance and short generation intervals, or slow viral reproduction and long generation intervals.

@@

Assuming that the intrinsic generation-interval distributions for asymptomatic and symptomatic cases follow gamma distributions with different means, $\bar G_a$ and  $\bar G_s$, and dispersions (represented by squared coefficients of variations), $\kappa_a$ and $\kappa_s$. 
Then, we have
\begin{equation}
\frac{q}{1-q}=\left(\frac{z}{1-z}\right)\left[\frac{(1 + \kappa_s r \bar G_s)^{1/\kappa_s}}{(1 + \kappa_a r \bar G_a)^{1/\kappa_a}}\right].
\label{eq.gammaratio}
\end{equation}
As is apparent from Eq.~\req{eq.qratio}, relevance $q$ increases with $z$, i.e., with both $p$ and $\mathcal R_a/\mathcal R_s$. The consequences for outbreak potential further depends on the relative lengths of generation intervals as well as the relative amount of dispersions in them.

%The basic reproduction number, ${\cal{R}}_0$,
%characterize the strength of an epidemic outbreak in an other
%wise susceptible population:
%\begin{equation}
%{\cal{R}}_{0}= p\left(\frac{\beta_a}{\gamma_a}\right)+ (1-p)\left(\frac{\beta_s}{\gamma_s}\right)
%\end{equation}
%Pooling estimates via a Bayesian approach applied to the early stages of Wuhan case data,
%suggests ${\cal{R}}_0\approx 3.0$ 
%with a 95\% confidence interval from 2-4.5.  Yet, that information
%is not sufficient to disentangle the role of asymptomatic cases, i.e.,
%what fraction of secondary cases can be ascribed to transmission
%from asymptomatic cases vs.~symptomatic cases?  A step towards such
%a link is enabled given the relationship between strength and speed, $r$,
%i.e., the rate at which cases grow in number and which we assume
%should be the same for both classes of cases.  Strength and size
%can be related via a Gamma approximation as:
%\begin{equation}
%{\cal{R}}^{gamma}=(1+\kappa r \bar{G})^{1/\kappa}
%\end{equation}
%where $\bar{G}$ is the mean generation interval and
%$\kappa$  is the squared coefficient of variation of
%the generation interval.  This approximation (explored in Champredon
%et al.) reveals why for a given observed speed,
%increases in the mean generation interval imply
%larger reproduction number. Reports of individuals
%with mild cases that remain infectious for (relatively)
%long periods suggest that the mean generation interval
%could be longer than estimated from symptomatic cases alone - both
%increasing ${\cal{R}}_0$ and driving more secondary cases
%than expected. Yet the relevance of asymptomatic cases
%depends on their realized proportion at a population level.

%The number of infected cases is expected
%to increase exponentially at rate $r$ at the outset of an outbreak.  We note
%that, in practice,
%counted cases may largely be of the symptomatic type. 
%Hence, we ask: is the expected ratio of $I_a/I_s$ 
%during the outbreak phase the same or different
%than $p/(1-p)$?  In other words, is the ratio of asymptomatic to symptomatic cases population-wide approximately equal to the ratio of asymptomatic to symptomatic cases generated by newly exposed individuals?
%To answer this question, note that the equations
%for the infectious cases can be rewritten given the ansatz
%$E(t)=c_e e^{rt}$,
%$I_a(t)=c_a e^{rt}$,
%$I_s(t)=c_s e^{rt}$,
%given the prefactors $c_e$, $c_a$, and $c_s$
%such that
%\begin{eqnarray}
%r c_a&=&p\gamma_e c_e-\gamma_a c_a,\\
%r c_s&=&(1-p)\gamma_e c_e-\gamma_s c_s,
%\end{eqnarray} 
%which implies that
%\begin{eqnarray}
%c_a(r+\gamma_a)&=&p\gamma_e c_e,\\
%c_s(r+\gamma_s)&=&(1-p)\gamma_e c_e.
%\end{eqnarray} 
%Taking the ratio of these two equations yieldds
%\begin{equation}
%\frac{c_a}{c_s}=\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right].
%\label{eq.relevance}
%\end{equation}
%Alternatively, this same equation can be written as:
%\begin{equation}
%\frac{c_a}{c_s}=\frac{p}{1-p}\left[\frac{1+\tau/T_s}{1+\tau/T_a}\right]
%\end{equation}
%where $\tau$ is the characteristic time of exponential case increase,
%$T_s$ is the average period of symptomatic infectious periods,
%and $T_a$ is the average period of asymptomatic infectious periods. 
%Hence, we conclude that, during the exponential
%growth phase, the intrinsic
%ratio of asymptomatic to symptomatic cases ($p/(1-p)$) may differ from
%the realized ratio at the population level ($c_a/c_s$).  
%%This difference
%suggests that additional steps are required to link asymptomatic features
%at the individual to the population scale, i.e., by including
%the unknown transmission rates, $\beta_a$ and $\beta_s$.
%Indeed, the ratio of secondary case production caused by asymptomatic vs.~symptomatic
%individuals during the exponential phase should be
%\begin{equation}
%\frac{q}{1-q} = \left(\frac{\beta_a}{\beta_s}\right)\frac{p}{1-p}\left[\frac{r+\gamma_s}{r+\gamma_a}\right],
%\label{eq.qratio}
%\end{equation}
%where $q$ is the fraction of new secondary cases caused by
%asymptomatic individuals or `relevance' of asymptomatic cases.
%The value of $q$ is an unknown feature of
%outbreaks at the population scale, distinct from $p$ the probability of becoming
%asymptomatic upon infection. 
%As is apparent from Eq.~\req{eq.qratio}, relevance increases
%with both $p$ and $\beta_a/\beta_s$. However, the consequences
%for outbreak potential depends on the relative duration of infectiousness.

Here, given $1/r=7$ days, $\bar G_s=7$ days, and $\kappa_s=0.5$, we evaluate the impacts of intrinsic proportion of asymptomatic transmission $z$ on the realized relevance of asymptomatic infection, given variation in relative mean generation interval $\bar G_a/\bar G_s$ and relative generation-interval dispersion $\kappa_a/\kappa_s$ of asymptomatic cases.
We vary $\bar G_a/\bar G_s$ and $\kappa_a/\kappa_s$ between 0.5 and 2 one at a time, and infer $q$ and ${\cal{R}}_0$ compatible with an estimated value of $r$ from case data given the generation interval distributions of asymptomatic and asymptomatic cases.

@@

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the relevance of asymptomatic transmission and basic reproduction number, given variation in
the mean generation interval of asymptomatic cases. 
(A) Long/short generation intervals of asymptomatic cases decrease/increase the relevance of asymptomatic cases, $q$.
(B) Long/short generation intervals of asymptomatic cases increase/decrease the basic reproduction number ${\cal{R}}_0$.
(C) Wide/narrow generation intervals of asymptomatic cases increase/decrease the relevance of asymptomatic cases, $q$.
(D) Wide/narrow generation intervals of asymptomatic cases decrease/increase the basic reproduction number ${\cal{R}}_0$.
Here, we vary relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, while assuming $\kappa_a = \kappa_s$.
As our baseline parameters, we assume $1/r=7$ days, $\bar G_s=7$ days, and $\kappa_s=0.5$.
\label{fig.relevance}}
\end{center}
\end{figure*}

Across the ranges of parameters we explore, the intrinsic proportion of asymptomatic transmission $z$ is similar to the realized proportion $q$ (Figure~\ref{fig.relevance}A,C).
As the relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, increases, $q$ decreases because symptomatic cases are more likely to have short generation intervals (i.e., fast transmission events), which drive the spread during the growth phase (Figure~\ref{fig.relevance}A).
On the other hand, as the relative generation-interval dispersion of asymptomatic cases, $\kappa_a/\kappa_s$, increases, $q$ also increases because the amount of short asymptomatic generation intervals (i.e., the proportion of faster transmission by asymptomatic individuals) increases (Figure~\ref{fig.relevance}C).

Figure~\ref{fig.relevance}B compares the effect of relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, on ${\cal{R}}_0$.
When $\bar G_s < \bar G_a$, then the strength ${\cal{R}}_0$ increases with $z$.
This increase arises because asymptomatic infections that have long generation intervals increase the duration of overall generation intervals (see~\citep{park_2019practical})
Conversely, when $\bar G_s > \bar G_a$ then the strength ${\cal{R}}_0$ decreases with $z$ because short generation intervals of asymptomatic cases reduce the overall mean generation interval.

Likewise, Figure~\ref{fig.relevance}D compares the effect of relative generation-interval dispersion of asymptomatic cases, $\kappa_a/\kappa_s$, on ${\cal{R}}_0$.
Depending on whether $\kappa_a$ is larger or smaller than $\kappa_s$, increase in the proportion of asymptomatic transmission $z$ can either decrease or increase the strength ${\cal{R}}_0$.
In other words, when ${\cal{R}}_0$ is estimated without explicitly accounting for asymptomatic spread, it can be over- or under- estimated depending on the relative mean and dispersion in generation intervals of asymptomatic and symptomatic individuals.

%Here, given $1/r=7$ days, $1/\gamma_e=4$ days, and
%$1/\gamma_s=10$ days, we evaluate the impacts of $p$ and $\beta_a/\beta_s$ on the
%relevance of asymptomatic infections, given variation in 
%$p$, $\beta_a/\beta_s$, and $\gamma_a$.  We consider two
%scenarios, one in which asymptomatic infectiousness is short, i.e., with
%average duration of $1/\gamma_a=5$ days,
%and one in which asymptomatic infectiousness is long,i.e., $1/\gamma_a=20$ days.
%We use the renewal equation formalism to explicitly link duration of infectiousness to the 
%basic reproductive number (for technical details and example 
%applications see~\citep{wallinga2007generation,weitz_scirep2015,champredon_2018}).
%In brief, we
%infer ${\cal{R}}_0$ compatible with an estimated value of $r$ from case
%data given the generation interval distribution 
%(set by $q$ and the duration in the exposed and infectious states).
%By solving for $q$, we can then relate the basic
%reproduction number to the relative transmission rates ($\beta_a/\beta_s$)
%and the probability, $p$, of asymptomatic cases at the individual scale.
%Figure~\ref{fig.relevance}C shows that
%when $\gamma_a<\gamma_s$, then the strength ${\cal{R}}_0$ increases
%with $q$. 
%This increase arises because asymptomatic infections that resolve
%slower increase the duration of
%overall generation intervals (see~\citep{park_2019practical}). In contrast,
%Figure~\ref{fig.relevance}D show that
%when $\gamma_a>\gamma_s$ then
%the strength ${\cal{R}}_0$ decreases with $q$.  
%This decrease arises because asymptomatic infections that resolve
%faster decrease the duration of
%overall generation intervals (again, see~\citep{park_2019practical}). In both 
%cases, the dependence of ${\cal{R}}_0$ on $q$ is
%compared to a baseline of 
%${\cal{R}}_0\approx 3.1$ in a model with no secondary
%case production from asymptomatic infections (see filled
%circle at bottom left of Figure~\ref{fig.relevance}C-D.
%In other words, when ${\cal{R}}_0$ is estimated without explicitly accounting for asymptomatic spread, it can be over- or under- estimated depending on the relative duration of infection between symptomatic and asymptomatic individuals.

>>>>

In an epidemic, symptomatic cases are the predominant focus of treatment and usually represent the bulk of reported cases. 
However, infected individuals who are asymptomatic yet infectious can be a critical factor in the spread of some pathogens \citep{fraser2004factors}.
Asymptomatic individuals are hard to trace, unlikely to self-isolate, and are likely to retain normal social and travel patterns.
The COVID-19 outbreak has raised significant
questions regarding the role of asymptomatic cases~\citep{fauci_nejm2020}.
A key focus has been on estimating asymptomatic prevalence (see~\citep{mizumoto_2020}), e.g., what fraction of individuals have asymptomatic infections relative to those with symptomatic infections?

@@

We model viral spread using a renewal-equation framework \citep{heesterbeek1996concept}, which allows us to model the current incidence of infected individuals as a function of previous incidence and how infectiousness of an infected individual varies over the course of their infection.
We divide incidence $i$ into two categories -- $i_a$ and $i_s$ -- corresponding to incidence of asymptomatic and symptomatic cases, respectively.
Both types of individuals can infect others, but they differ in their intrinsic reproduction numbers, ${\cal{R}}_a$ and ${\cal{R}}_s$, as well as intrinsic generation-interval distributions, $g_a(\tau)$ and $g_s(\tau)$.
Generation intervals, which are defined as the time between when an individual is infected and when that individual infects another person \citep{svensson2007note}, depend on the natural history of infection:
individuals with subclinical cases may have fast clearance and short generation intervals, or slow viral reproduction and long generation intervals.
The shape of the generation-interval distribution characterizes the relationship between the epidemic growth rate $r$ and the reproduction number \citep{wallinga2007generation}.

@@

In order to understand how the shape of asymptomatic and symptomatic generation-interval distributions affects $q$ and $\mathcal R_0$, we summarize the distributions in terms of their means and dispersions.
Assuming that the intrinsic generation-interval distributions for asymptomatic and symptomatic cases follow gamma distributions with different means, $\bar G_a$ and  $\bar G_s$, and dispersions (represented by squared coefficients of variations), $\kappa_a$ and $\kappa_s$.
Then, we have
\begin{equation}
\frac{q}{1-q}=\left(\frac{z}{1-z}\right)\left[\frac{(1 + \kappa_s r \bar G_s)^{1/\kappa_s}}{(1 + \kappa_a r \bar G_a)^{1/\kappa_a}}\right].
\label{eq.gammaratio}
\end{equation}
As is apparent from Eq.~\req{eq.qratio}, relevance $q$ increases with $z$, i.e., with both $p$ and $\mathcal R_a/\mathcal R_s$. The consequences for outbreak potential further depends on the relative lengths of generation intervals as well as the relative amount of dispersions in them.
In Supplementary Materials, we use a compartmental ordinary differential equations model to show that the duration of infectious period of asymptomatic individuals affects both $q$ and $\mathcal R_0$, which is a special case of Eq.~\req{eq.qratio}.

@@

\begin{figure*}[b!]
\begin{center}
\includegraphics[width=0.4\textwidth]{figheatmap.pdf}
\mbox{\hspace{0.05\textwidth}}
\includegraphics[width=0.4\textwidth]{figheatmap_R0.pdf}
\caption{Effects of intrinsic proportion of asymptomatic transmission on the relevance of asymptomatic transmission and basic reproduction number, given variation in
the mean generation interval of asymptomatic cases. 
(A) Long/short generation intervals of asymptomatic cases decrease/increase the relevance of asymptomatic cases, $q$.
(B) Long/short generation intervals of asymptomatic cases increase/decrease the basic reproduction number ${\cal{R}}_0$.
As our baseline parameters, we assume $1/r=7$ days, $\bar G_s=7$ days, and $\kappa_s=\kappa_a=0.5$.
\label{fig.relevance}}
\end{center}
\end{figure*}

Here, given $1/r=7$ days, $\bar G_s=7$ days, and $\kappa_s=\kappa_a=0.5$, we evaluate the impacts of intrinsic proportion of asymptomatic transmission $z$ on the realized relevance of asymptomatic infection, given variation in relative mean generation interval $\bar G_a/\bar G_s$ of asymptomatic cases.
We vary $\bar G_a/\bar G_s$ between 0.5 and 2 and infer $q$ and ${\cal{R}}_0$ compatible with an estimated value of $r$ from case data given the generation interval distributions of asymptomatic and asymptomatic cases.
Across the ranges of parameters we explore, the intrinsic proportion of asymptomatic transmission $z$ is similar to the realized proportion $q$ (Figure~\ref{fig.relevance}A).
As the relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, increases, $q$ decreases because symptomatic cases are more likely to have short generation intervals (i.e., fast transmission events), which drive the spread during the growth phase (Figure~\ref{fig.relevance}A).

Figure~\ref{fig.relevance}B compares the effect of relative mean generation interval of asymptomatic cases, $\bar G_a/\bar G_s$, on ${\cal{R}}_0$.
When $\bar G_s < \bar G_a$, then the strength ${\cal{R}}_0$ increases with $z$.
This increase arises because asymptomatic infections that have long generation intervals increase the duration of overall generation intervals (see~\citep{park_2019practical})
Conversely, when $\bar G_s > \bar G_a$ then the strength ${\cal{R}}_0$ decreases with $z$ because short generation intervals of asymptomatic cases reduce the overall mean generation interval.
The qualitative effects of $z$ and $\bar G_a/\bar G_s$ on $q$ and ${\cal{R}}_0$ remain robust when we assume narrower ($\kappa_s = \kappa_a = 0.3$; Figure~S1) or wider ($\kappa_s = \kappa_a = 0.8$; Figure~S2) generation intervals.

Relative generation-interval dispersion of asymptomatic cases $\kappa_a/\kappa_s$ also have qualitatively similar effects on $q$ and ${\cal{R}}_0$.
When the asymptomatic cases have a wider/narrower generation-interval distribution, higher intrinsic proportion of asymptomatic transmission $z$ increases/decreases the relevance of asymptomatic transmission $q$ but decreases/increases the basic reproduction number (Figure~S3).
Since a wider generation-interval distribution has a higher proportion of early transmission, it has qualitatively similar effects on $q$ and ${\cal{R}}_0$ as a generation-interval distribution with a small mean.
